Neural Network to Vatiational Sparse Gaussian Process: For Adaptive Exploration

### 2023-04-18 paper/main.tex
Aidan updated abstract

\begin{abstract}
Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
\end{abstract}

>>>>

\begin{abstract}
% Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
%
  Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning cannot adapt to new data because it relies upon gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data without further training. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

  % Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment
  % in real-world applications, e.g. healthcare and robotics.
  % % The ability to adapt to new information and express uncertainty is essential for deploying learning-based systems in real-world applications, e.g. healthcare and robotics.
  % However, modern deep learning cannot adapt to new data because it relies upon gradient-based optimisation,
  % which requires retraining with the addition of new data.
  % In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data without further training.
  % % Although, traditional approaches are limited to low-dimensional and small data sets, recent advances have scaled
  % % GPs to large data sets.
  % However, they are limited to low-dimensional and small data sets.
  % % We present a method that enables any trained neural network to adapt to new data without further gradient-based optimisation,
  % % whilst providing scalable and principled uncertainty quantification.
  % We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian
  % process (SVGP).
  % % We present a method that enables any trained neural network to adapt to new data without further gradient-based optimisation,
  % % whilst providing scalable and principled uncertainty quantification.
  % % Our method u converts a trained neural network to a sparse variational Gaussian process (SVGP).
  % Importantly, our method 1) enables fast adaptation to new data,
  % 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets.
  % % Importantly, our method 1) enables fast adaptation to new data,
  % % 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets.
  % Essentially, our method reaps the benefits of nonparametric models without their well-known issues.
  % % 2) does not require further gradient-based optimisation,
  % We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in
  % model-based reinforcement learning and
  % we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

  % obtains principled uncertainty estimates by
  % Our experiments demonstrate that our model's adaptivity can improve the sample-efficiency of model-based reinforcement
  % learning.

  % in model-based reinforcement learning using updated information for more efficient exploration.

  % Our method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available.

  % % While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch.
  % Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights.
  % Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process.
  % The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining.
  % Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available.
  % We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
\end{abstract}
